---
title: "ESM204_HW3"
author: "Camila Bobroff & Robert Heim"
date: "5/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, load libraries}
library(tidyverse)
```

```{r, read in the data}
data <- read_csv("HW3Data.csv")
```

Part 1. Create a linear probability model that predicts a respondent's probability of voting 'yes' on the ballot based on their age, income, NEP score, the risk reduction offered by the program, and the cost of the program to that respondent. Show the model and interpret the regression coefficients.

```{r}

vsr_lm <- lm(vote ~ risk + bid + NEP + income + age, data = data)
summary(vsr_lm)

intercept <- vsr_lm$coefficients[1]
risk <- vsr_lm$coefficients[2]
bid <- vsr_lm$coefficients[3]
nep <- vsr_lm$coefficients[4]
income_one <- vsr_lm$coefficients[5]
income_poor <- vsr_lm$coefficients[6]
income_rich <- vsr_lm$coefficients[7]
income_veryrich <- vsr_lm$coefficients[8]
age_50 <- vsr_lm$coefficients[9]
age_40 <- vsr_lm$coefficients[10]
age_60 <- vsr_lm$coefficients[11]
age_30 <- vsr_lm$coefficients[12]

```

Regression model:
$Pr(Voting~Yes) = `r round(intercept, digits = 4)` + `r round(age_30, digits=4)`(Age~to~30) + `r round(age_40, digits=4)`(Age~to~40) + `r round(age_50, digits=4)`(Age~to~50) + `r round(age_60, digits=4)`(Age~to~60) + `r round(income_one, digits=4)`(Income~One~Percent) + `r round(income_poor, digits=4)`(Income~Poor) + `r round(income_rich, digits=4)`(Income~Rich) + `r round(income_veryrich,digits=4)`(Income~Very~Rich) + `r round(nep, digits=4)`(NEP) + `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(Risk~Reduction)$

Interpretation of Regression coefficients:
- Risk:
- Bid:
- NEP:
Income reference level is middle
- Income one percent:
- Income poor:
- Income rich:
- Income very rich:
Age reference level is over 60
- to 30:
- to 40:
- to 50:
- to 60:

Part 2. Based on the regression, what can you say about the value of a single prevented whale death? (Hint: Think about how risk reduction for all whales translates into the number of whale deaths avoided)

Risk reduction (vessel speed reduction) = 20% -> 5 whales saved, given that ratio is constant:
Risk reduction (vessel speed reduction) = 4% -> 1 whale saved. For value of a single whale, find willingness to pay for 0% risk reduction and compare to 4% risk reduction
```{r}
# Assume the average probability of voting yes is the average of the votes
avg_vote <- mean(data$vote) # 0.714
avg_nep <- mean(data$NEP) # 38.366
# Assume age = to thirty and income = middle

# find bid for 0% risk reduction
# QUESTION: WHY ARE THE BIDS NEGATIVE? 

bid_0 <- (intercept + age_30 + income_rich + (nep*avg_nep) + (risk*0) - avg_vote)/bid
# Bid_0 = -39.5

bid_4 <- (intercept + age_30 + income_rich + (nep*avg_nep) + (risk*4) - avg_vote)/bid
# Bid_4 = -42.3

# The value of a single prevented whale death is difference between willingness to pay of 4% and 0% vessel speed reduction

single_whale <- bid_0 - bid_4
# 2.78

```

Part 3. Pick three arbitrary respondents. For each of those respondents, estimate their WTP for a VSR program offering 60% risk reduction.

```{r}
# Randomly select three arbitrary respondents 
sample(1:500, 3, replace = FALSE) # 372, 362, 269

vsr_random <- data %>% # dataframe with only those 3 random respondents
  slice(372, 362, 269)
# risk = 40, bid = 80, vote = 0, NEP = 48, income = very rich, age = to 50
# risk = 60, bid = 80, vote = 1, NEP = 47, income = middle, age = over 60
# risk = 80, bid = 20, vote = 1, NEP = 32, income = very rich, age = to 40

# Estimate WTP at 60% risk reduction
# QUESTION: WHY ARE SOME OF THE BIDS NEGATIVE? 

# Person 372
bid_372 <- (intercept + age_50 + income_veryrich + (nep*48) + (risk*60) - avg_vote)/bid
# -251

# Person 362
bid_362 <- (intercept + age_60 + (nep*47) + (risk*60) - avg_vote)/bid
# -168

# Person 269
bid_269 <- (intercept + age_40 + (nep*32) + (risk*60) - avg_vote)/bid
# 58.1
```

Part 4. Now repeat this process and estimate the mean WTP for a VSR program offering 60% risk reduction among Santa Barbara County households

```{r}
# Use average age (tofifty), nep(38.366), vote(0.714), and income(rich) for Santa Barbara County

ave_income <- data %>% 
  count(income == "very_rich", income == "rich", income == "poor", income == "one_percent", income == "middle")
# Middle income is rich

ave_age <- data %>% 
  count(age == "tothirty", age == "toforty", age == "tofifty", age == "tosixty", age == "oversixty")
# To fifty is middle age

bid_sb <- (intercept + age_50 + (nep*38.366) + (risk*60) - avg_vote)/bid
# -64.5

```

Part 5. If there are 150,000 households in Santa Barbara County, estimate the total benefit of a VSR program to these households

```{r}
# Multiply bid_sb by 150,000

total_benefit <- bid_sb*150000
# -9,671,402

```

Part 6. Based on only these benefits, do the benefits of the VSR program outweigh the costs?

Total benefit to SB residents is 9,671,402, which outweighs the cost of implementing the VSR program (7 million). The net benefit of the program is 2,671,402.

