---
title: "ESM204_HW3"
author: "Camila Bobroff & Robert Heim"
date: "5/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, load libraries}
library(tidyverse)
```

```{r, read in the data}
data <- read_csv("HW3Data.csv")
```


Part 1. Create a linear probability model that predicts a respondent's probability of voting 'yes' on the ballot based on their age, income, NEP score, the risk reduction offered by the program, and the cost of the program to that respondent. Show the model and interpret the regression coefficients.
```{r}

vsr_lm <- lm(vote ~ risk + bid + NEP + income + age, data = data)
summary(vsr_lm)

intercept <- vsr_lm$coefficients[1]
risk <- vsr_lm$coefficients[2]
bid <- vsr_lm$coefficients[3]
nep <- vsr_lm$coefficients[4]
income_one <- vsr_lm$coefficients[5]
income_poor <- vsr_lm$coefficients[6]
income_rich <- vsr_lm$coefficients[7]
income_veryrich <- vsr_lm$coefficients[8]
age_50 <- vsr_lm$coefficients[9]
age_40 <- vsr_lm$coefficients[10]
age_60 <- vsr_lm$coefficients[11]
age_30 <- vsr_lm$coefficients[12]

```

Regression model:
$Pr(Voting~Yes) = `r round(intercept, digits = 4)` + `r round(age_30, digits=4)`(Age~to~30) + `r round(age_40, digits=4)`(Age~to~40) + `r round(age_50, digits=4)`(Age~to~50) + `r round(age_60, digits=4)`(Age~to~60) + `r round(income_one, digits=4)`(Income~One~Percent) + `r round(income_poor, digits=4)`(Income~Poor) + `r round(income_rich, digits=4)`(Income~Rich) + `r round(income_veryrich,digits=4)`(Income~Very~Rich) + `r round(nep, digits=4)`(NEP) + `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(Risk~Reduction)$

Interpretation of Regression coefficients:
- Risk:
- Bid:
- NEP:
Income reference level is middle
- Income one percent:
- Income poor:
- Income rich:
- Income very rich:
Age refernce level is over 60
- to 30:
- to 40:
- to 50:
- to 60:

Part 2. Based on the regression, what can you say about the value of a single prevented whale death? (Hint: Think about how risk reduction for all whales translates into the number of whale deaths avoided)

Risk reduction (vessel speed reduction) = 20% -> 5 whales saved, given that ratio is constant:
Risk reduction (vessel speed reduction) = 4% -> 1 whale saved. For value of a single whale, find willingness to pay for 0% risk reduction and compare to 4% risk reduction
```{r}



```

